MADA Project - Ibis Blood Parasites from Florida

### 2021-10-29 products/manuscript/Manuscript.Rmd
Completed Project Part 3 10/29/21
Completed part 3 of the MADA project. Statistical analysis scripts contains code to create 10 models, and save them. Analyses afterwards will most likely focus on the presence absence data, and work on tweaking the GLM models already in place.

save_data_location3 <- here::here("data", "processed_data", "processedIbisFielddata.rds")
saveRDS(IbisField, file = save_data_location3)
```

@@

################################Saving the UrbanIbis Data#############################
#Saving the Ibis Blood 15-17 dataset as an RDS in the processed data folder
save_data_location3 <- here::here("data", "processed_data", "processedIbisUrban.rds")
saveRDS(IbisUrban, file = save_data_location3)

@@

Figure_file1 <- here("results", "ExploratoryAnalysis", "ExpFigure1.png")
ggsave(filename = Figure_file1,plot = ExpFigure1)

@@

Figure_file2 <- here("results", "ExploratoryAnalysis", "ExpFigure2.png")
ggsave(filename = Figure_file2, plot = ExpFigure2)

@@

Figure_file3 <- here("results", "ExploratoryAnalysis", "ExpFigure3.png")
ggsave(filename = Figure_file3, plot = ExpFigure3)

@@

Figure_file4 <- here("results", "ExploratoryAnalysis", "ExpFigure4.png")
ggsave(filename = Figure_file4, plot = ExpFigure4)

@@

Figure_file5 <- here("results", "ExploratoryAnalysis", "ExpFigure5.png")
ggsave(filename = Figure_file5, plot = ExpFigure5)

@@

Figure_file6 <- here("results", "ExploratoryAnalysis", "ExpFigure6.png")
ggsave(filename = Figure_file6, plot = ExpFigure6)

@@

>>>>

save_data_location3 <- here::here("processed_data", "processedIbisFielddata.rds")
saveRDS(IbisField, file = save_data_location3)
```

@@

################################Saving the UrbanIbis Data#############################
#Saving the Ibis Blood 15-17 dataset as an RDS in the processed data folder
save_data_location3 <- here::here("processed_data", "processedIbisUrban.rds")
saveRDS(IbisUrban, file = save_data_location3)

@@

Figure_file1 <- here("results", "Figures", "ExpFigure1.png")
ggsave(filename = Figure_file1,plot = ExpFigure1)

@@

Figure_file2 <- here("results", "Figures", "ExpFigure2.png")
ggsave(filename = Figure_file2, plot = ExpFigure2)

@@

Figure_file3 <- here("results", "Figures", "ExpFigure3.png")
ggsave(filename = Figure_file3, plot = ExpFigure3)

@@

Figure_file4 <- here("results", "Figures", "ExpFigure4.png")
ggsave(filename = Figure_file4, plot = ExpFigure4)

@@

Figure_file5 <- here("results", "Figures", "ExpFigure5.png")
ggsave(filename = Figure_file5, plot = ExpFigure5)

@@

Figure_file6 <- here("results", "Figures", "ExpFigure6.png")
ggsave(filename = Figure_file6, plot = ExpFigure6)

@@

## Statistical Analyses

### Presence/Absence Haemoproteus Parasitemia Data

This section details the results of a preliminary statistical analysis on the Haemoproteus parasitemia data just looking at the presence and absence data, as opposed to the amount of parasitemia per sampled bird.

The following models were setup and tested through a 10-fold cross-validation. The training/testing data proportions were 80/20. 10 folds were taken on 80% of the observations, and the results are depicted here. I am currently struggling with applying the models generated by CV to the testing data.

```{r}
data_locationPA1 <- here::here("results", "Tables", "Model1MetricsPA.Rda")
table1 <- readRDS(data_locationPA1)
table1
```
This univariate classification GLM looked at the effect of habitat type on the presence of Haemoproteus in infected birds. The results were unsatisfactory.

```{r}
data_locationPA2 <- here::here("results", "Tables", "Model2MetricsPA.Rda")
table2 <- readRDS(data_locationPA2)
table2
```
This univariate classification GLM looked at the effect of sex of the Ibis on the presence of Haemoproteus in infected birds. The results were unsatisfactory.

```{r}
data_locationPA3 <- here::here("results", "Tables", "Model3MetricsPA.Rda")
table3 <- readRDS(data_locationPA3)
table3
```
This univariate classification GLM looked at the effect of age of the Ibis on the presence of Haemoproteus in infected birds. The results were unsatisfactory.

```{r}
data_locationPA4 <- here::here("results", "Tables", "Model4MetricsPA.Rda")
table4 <- readRDS(data_locationPA4)
table4
```
This univariate classification GLM looked at the effect of sampling site on the presence of Haemoproteus in infected birds. The results were unsatisfactory.

```{r}
data_locationPA5 <- here::here("results", "Tables", "Model5MetricsPA.Rda")
table5 <- readRDS(data_locationPA5)
table5
```
This univariate classification GLM looked at the effect of sampling date on the presence of Haemoproteus in infected birds. The results were unsatisfactory.

```{r}
data_locationPA6 <- here::here("results", "Tables", "Model6MetricsPA.Rda")
table6 <- readRDS(data_locationPA6)
table6
```
This univariate classification GLM looked at the effect of sampling season on the presence of Haemoproteus in infected birds. The results were unsatisfactory, however, these results were better than the other univariate models so far.

```{r}
data_locationPA7 <- here::here("results", "Tables", "Model7MetricsPA.Rda")
table7 <- readRDS(data_locationPA7)
table7
```
This multivariate classification GLM looked at the effect of age and sex on the presence of Haemoproteus in infected birds. The results were unsatisfactory, this model scored about the same as the previous univariate models for age and sex.

```{r}
data_locationPA8 <- here::here("results", "Tables", "Model8MetricsPA.Rda")
table8 <- readRDS(data_locationPA8)
table8
```
This multivariate classification GLM looked at the effect of sampling date and sampling season on the presence of Haemoproteus in infected birds. The results here had the highest performance of all the other models so far. One way to interpret this model would be that as the sampling effort continued from 2010-2017, the passage of time and season had a significant effect on the presence of Haemoproteus in Ibis. However, the performance of this model isn't high enough to successfully conclude with that explanation.

### Regression Models for the Amount of Haemoproteus Parasitemia

This section details the results of a preliminary statistical analysis on the Haemoproteus parasitemia data looking at the amount of measured parasitemia per sampled bird (on a logarithm base 10 scale), as opposed to presence/absence data.

The following models were setup and tested through a 10-fold cross-validation. The training/testing data proportions were 80/20. 10 folds were taken on 80% of the observations, and the results are depicted here. I am currently struggling with applying the models generated by CV to the testing data.

#### Random Forest Model

The following are the results from a random forest model run on all the predictors against the Haemoproteus parasitemia (log base 10) data. The missing data values were imputed by K nearest neighbors. The model was run on all 10 folds from the training dataset.
```{r}
data_locationRF <- here::here("results", "Tables", "RandomForestMetrics.Rda")
table9 <- readRDS(data_locationRF)
table9
```
This random forest model looked at the effect of all predictors on the amount of Haemoproteus in infected birds. The results were unsatisfactory, this model had a relatively high RMSE and low R squared value.

#### Linear Regression Model

The following are the results from a linear regression model run on habitat type against the Haemoproteus parasitemia (log base 10) data.

```{r}
data_locationLM1 <- here::here("results", "Tables", "LinMetrics1.Rda")
table10 <- readRDS(data_locationLM1)
table10
```
This linear regression model looked at the effect of habitat type on the amount of Haemoproteus in infected birds. The results were unsatisfactory, the RMSE is high and the R squared value is very low.

#### Conclusions so far

This concludes the statistical analysis for part 3. Overall, the models did not perform well, especially those looking at the amount of Haemoproteus in infected birds. The presence/absence data has more potential for conclusions to be made based on generated models. The next steps would be to fine-tune the classification models for the presence/absence data and create a random forest model.

Current problems that have arose in this process include imputing missing values to run models (as there is a large amount of missing data in this dataset), and visualizing the tidymodels results on a graph.
